### 2016-09-16 paper/implementation.tex
added rough draft of all sections

\subsection{Synthesis Overview}
\begin{figure}
\includegraphics[scale=.5]{synth-lens-schematic.png}
\label{fig:synth-lens-schematic}
\end{figure}
The synthesis algorithm is shown in a schematic diagram in 
Figure~\ref{fig:synth-lens-schematic}.  The algorithm takes in a pair of
regular expressions, and examples, and converts the regular expressions into
DNF regular expressions, using \ToDNFRegex{}.
These then get enqueued, and SynthDNFLens, where the bulk of the work
occurs.

@@

$\Distance_{dist}$ captures how far apart the distribution of
user defined regular expressions are.
The distribution of user defined regular expressions can be viewed as an
infinite dimensional free module over the integers, \Module{}.
The basis for this is $\SetOf{1*(\RegexVariable,n) | n\in\Nats, \RegexVariable
\text{ is a user defined regular expression}}$
DNF regular expressions with variables can map into this with a function by
counting the number of user defined data types present at a given level.
The formal definition for this mapping is given by the function \GetDist{}.

\begin{definition}\leavevmode\\
\label{def:getdist}
\begin{tabular}{@{}L@{}L@{}}
\GetDist(\DNFOf{\Sequence_1;\ldots;\Sequence_n}) &
=\GetDist(\Sequence_1)+\ldots+\GetDist(\Sequence_n)\\
\GetDist(\SequenceOf{\String_0;\Atom_1;\ldots;\Atom_n;\String_n}) &
=\GetDist(\Atom_1)+\ldots+\GetDist(\Atom_n)\\
\GetDist(\RegexVariable)&=1*(\RegexVariable,1)\\
\GetDist(\IterateLens{\DNFLens})&=\phi(\GetDist(\DNFLens))
\end{tabular}

@@

$\Distance_{dist}(\DNFRegex,\DNFRegexAlt)=\LOneNorm(\phi(\DNFRegex),\phi(\DNFRegexAlt))$
where \LOneNorm{} is the taxicab metric on \Module{}.
The priority of a pair of DNF regular expressions, $(\DNFRegex{},\DNFRegexAlt{})$,
is $\Distance(\DNFRegex{},\DNFRegexAlt{})+8*expansion_{count}$.
Keeping track of expansion count is important as it allows for lower priority
expansions to be eventually popped, instead of getting stuck doing incorrect,
increasingly deep expansions whose DNF regular expressions have a low distance.
\afm{should i do an example, or is that going into too much detail?}
Experimentally, 8 was a good value for how much weight the expansion
count should get in the priority.  It allows for not getting stuck in a local
minima in the search space, while still allowing for the distance to choose the
correct expansion the majority of the time.

@@

\begin{definition}
  Given an ordering $\leq : A \rightarrow A \rightarrow \SetOf{-1,0,1}$,
  and $\ListOf{a_1;\ldots;a_n}\in A \ListType$
  We define $\SortingOf{\leq}{\ListOf{a_1;\ldots;a_n}}$ as
  $\sigma \in \PermutationSetOf{n}$ where if $i\leq j$,
  then $a_{\sigma(i)} \leq a_{\sigma(j)}$.

@@

  \item
    $\StarOf{\DNFRegex} \AtomLeq \StarOf{\DNFRegexAlt}$
    if
    $\DNFRegex \DNFLeq \DNFRegexAlt$
  \end{enumerate}
\end{definition}

@@

\begin{definition}
  \begin{enumerate}
  \item A DNF regular expression $\DNFOf{Sequence_1;\ldots;\Sequence_n}$ is
    \textit{normalized} if $i \leq j$ implies $\Sequence_1\SequenceLeq\Sequence_j$,
    and each $\Sequence_i$ is normalized.
  \item A sequence $\SequenceOf{\Atom_1;\ldots;\Atom_n}$ is \textit{normalized}
    if $i \leq j$ implies $\Atom_i\AtomLeq\Atom_j$,
    and each $\Atom_i$ is normalized.
  \item An atom $\StarOf{\DNFRegex}$ is \textit{normalized} if
    $\DNFRegex$ is normalized.
  \end{enumerate}
\end{definition}

@@

\begin{definition}
  \begin{enumerate}
  \item $\NormalizeOf{\DNFOf{\Sequence_1;\ldots;\Sequence_n}} =
    \DNFOf{\NormalizeOf{\Sequence_{\sigma(1)}};\ldots;\NormalizeOf{\Sequence_{\sigma(n)}}}$,
    where
    $\sigma=\SortingOf{\SequenceLeq}
    {\ListOf{\NormalizeOf{\Sequence_1};\ldots;\NormalizeOf{\Sequence_n}}}$
  \item $\NormalizeOf{\SequenceOf{\String_0;\Atom_1;\ldots;\Atom_n;\String_n}} =
    \SequenceOf{\String_0;\Atom_{\sigma(1)};\ldots;\Atom_{\sigma(n)};\String_n}$,
    where
    $\sigma=\SortingOf{\SequenceLeq}
    {\ListOf{\NormalizeOf{\Atom_1};\ldots;\NormalizeOf{\Atom_n}}}$
  \item $\NormalizeOf{\StarOf{\DNFRegex}}=\StarOf{\NormalizeOf{\DNFRegex}}$
  \end{enumerate}
\end{definition}

Critically, the normalized form of the DNF regular expression, and the ordering
on DNF regular expressions, are tied to the existance of a DNF Lens without
rewrites.

\begin{theorem}
  $\NormalizeOf{\DNFRegex}\DNFLeq\NormalizeOf{\DNFRegexAlt}$ and
  $\NormalizeOf{\DNFRegexAlt}\DNFLeq\NormalizeOf{\DNFRegex}$ if, and only
  if, there exists a $\DNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt$,
  where $\DNFLens$ has no applications of a rewrite rule.
\end{theorem}

This can be proven with mutual induction, proving similar statements for
sequences and atoms.

Furthermore, two DNF regular expressions can be zipped together to form a
DNF lens.  Consider $\DNFOf{\Sequence_{\sigma_1(0)};\ldots;\Sequence_{\sigma_1(n)}}$, and
$\DNFOf{\SequenceAlt_{\sigma_2(0)};\ldots;\SequenceAlt_{\sigma_2(n)}}$, two normalized
DNF regular expressions.  $\Sequence_{\sigma(0)}$ merely needs to be mapped to
$\SequenceAlt_{\tau(0)}$.  Furthermore, the permutation mapping the two lenses
merely needs to be $\sigma_1 \Compose \sigma_2^{-1}$

We can use these orderings to find

Beyond that, we can use these orderings to find the lenses.
If $\DNFOf{\Sequence_1;\ldots;\Sequence_n}\leq\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}$,
and $\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}\leq\DNFOf{\Sequence_1;\ldots;\Sequence_n}$,
then $\Sorting([\Sequence_1;\ldots;\Sequence_n])\leq\Sorting([\Sequence_1;\ldots;\Sequence_n])$
using the dictionary ordering, which means that
$\Sorting([\Sequence_1;\ldots;\Sequence_n])[i]\leq
\Sorting([\SequenceAlt_1;\ldots;\SequenceAlt_n])[i]$ and
$\Sorting([\SequenceAlt_1;\ldots;\SequenceAlt_n])[i]\leq
\Sorting([\Sequence_1;\ldots;\Sequence_n])[i]$, which means that there is a
sequence lens between them.
\afm{I can keep going on about this stuff, should it go here?  Should I just make a brief claim
about it and move on}

Unfortunately, while this can find a well typed DNF regular expression, it doesn't
necessarily find the one that matches the examples.  In the situation where
there are multiple valid sorted orderings, there are multiple different lenses
with potentially different semantics.  Instead of iterating through all the
possible sortings, and finding one that matches the examples, we would like to
be able to immediately find only a sorting which satisfies the examples.

@@

This is done by joining additional parsing information alongside the regular
expression components.

>>>>

\subsection{Synthesis Overview}
\begin{figure}
\includegraphics[scale=.5]{synth-lens-schematic.png}
\label{fig:synth-lens-schematic}
\end{figure}
The synthesis algorithm is shown in a schematic diagram in 
Figure~\ref{fig:synth-lens-schematic}.  The algorithm takes in a pair of
regular expressions, and examples, and converts the regular expressions into
DNF regular expressions, using $\ToDNFRegex{}$.
These then get enqueued, and SynthDNFLens, where the bulk of the work
occurs.

@@

$\Distance_{dist}$ captures how far apart the distribution of
user defined regular expressions are.
The distribution of user defined regular expressions can be viewed as an
infinite dimensional vector space over the reals, $\VectorSpace{}$.
The basis for this is $\SetOf{1*(\RegexVariable,n) | n\in\Nats, \RegexVariable
\text{ is a user defined regular expression}}$
DNF regular expressions with variables can map into this with a function by
counting the number of user defined data types present at a given level.
The formal definition for this mapping is given by the function \GetDist{}.

\begin{definition}\leavevmode\\
\label{def:getdist}
\begin{tabular}{@{}L@{}L@{}}
\GetDist(\DNFOf{\Sequence_1;\ldots;\Sequence_n}) &
=\GetDist(\Sequence_1)+\ldots+\GetDist(\Sequence_n)\\
\GetDist(\SequenceOf{\String_0;\Atom_1;\ldots;\Atom_n;\String_n}) &
=\GetDist(\Atom_1)+\ldots+\GetDist(\Atom_n)\\
\GetDist(\RegexVariable)&=1*(\RegexVariable,1)\\
\GetDist(\IterateLensOf{\DNFLens})&=\phi(\GetDist(\DNFLens))
\end{tabular}

@@

$\Distance(\DNFRegex,\DNFRegexAlt)=\LOneNorm(\GetDist(\DNFRegex),\GetDist(\DNFRegexAlt))$
where $\LOneNorm$ is the taxicab metric on $\VectorSpace{}$.
The priority of a pair of DNF regular expressions, $(\DNFRegex{},\DNFRegexAlt{})$,
is $\Distance(\DNFRegex{},\DNFRegexAlt{})+8*expansion_{count}$.
Keeping track of expansion count is important as it allows for lower priority
expansions to be eventually popped, instead of getting stuck doing incorrect,
increasingly deep expansions whose DNF regular expressions have a low distance.
\afm{should i do an example, or is that going into too much detail?}
Experimentally, 8 was a good value for how much weight the expansion
count should get in the priority.  It allows for not getting stuck in a local
minima in the search space, while still allowing for the distance to choose the
correct expansion the majority of the time.

@@

\begin{definition}
  Given an ordering $\leq : A \rightarrow A \rightarrow \SetOf{-1,0,1}$,
  and $\ListOf{a_1;\ldots;a_n}\in \ListTypeOf{A}$
  We define $\SortingOf{\leq}{\ListOf{a_1;\ldots;a_n}}$ as
  $\sigma \in \PermutationSetOf{n}$ where if $i\leq j$,
  then $a_{\sigma(i)} \leq a_{\sigma(j)}$.

@@

  \item
    $\StarOf{\DNFRegex} \AtomLeq \StarOf{\DNFRegexAlt}$
    if
    $\DNFRegex \DNFLeq \DNFRegexAlt$

  \item 
    $\DNFRegex \DNFEq \DNFRegexAlt$
    if
    $\DNFRegex \DNFLeq \DNFRegexAlt$ and $\DNFRegexAlt \DNFLeq \DNFRegex$.

  \item
    $\Sequence \SequenceEq \SequenceAlt$
    if
    $\Sequence \SequenceLeq \SequenceAlt$ and $\SequenceAlt \SequenceLeq
    \Sequence$.

  \item
    $\Atom \AtomEq \AtomAlt$
    if
    $\Atom \AtomLeq \AtomAlt$ and $\AtomAlt \AtomLeq \Atom$.
  \end{enumerate}
\end{definition}

@@

\begin{definition}
  \begin{enumerate}
  \item A DNF regular expression $\DNFOf{\Sequence_1;\ldots;\Sequence_n}$ is
    \textit{normalized} if $i \leq j$ implies $\Sequence_1\SequenceLeq\Sequence_j$,
    and each $\Sequence_i$ is normalized.
  \item A sequence $\SequenceOf{\Atom_1;\ldots;\Atom_n}$ is \textit{normalized}
    if $i \leq j$ implies $\Atom_i\AtomLeq\Atom_j$,
    and each $\Atom_i$ is normalized.
  \item An atom $\StarOf{\DNFRegex}$ is \textit{normalized} if
    $\DNFRegex$ is normalized.
  \end{enumerate}
\end{definition}

@@

\begin{definition}
  \begin{enumerate}
  \item $\NormalizeOf{\DNFOf{\Sequence_1;\ldots;\Sequence_n}} =
    \DNFOf{\NormalizeOf{\Sequence_{\sigma(1)}};\ldots;\NormalizeOf{\Sequence_{\sigma(n)}}}$,
    where
    $\sigma=\SortingOf{\SequenceLeq}
    {\ListOf{\NormalizeOf{\Sequence_1};\ldots;\NormalizeOf{\Sequence_n}}}$
  \item $\NormalizeOf{\SequenceOf{\String_0;\Atom_1;\ldots;\Atom_n;\String_n}} =
    \SequenceOf{\String_0;\Atom_{\sigma(1)};\ldots;\Atom_{\sigma(n)};\String_n}$,
    where
    $\sigma=\SortingOf{\AtomLeq}
    {\ListOf{\NormalizeOf{\Atom_1};\ldots;\NormalizeOf{\Atom_n}}}$
  \item $\NormalizeOf{\StarOf{\DNFRegex}}=\StarOf{\NormalizeOf{\DNFRegex}}$
  \end{enumerate}
\end{definition}

\begin{theorem}
  \begin{enumerate}
  \item
    If $\DNFOf{\Sequence_1;\ldots;\Sequence_n}$ and
    $\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_m}$ are normalized, and
    $\DNFOf{\Sequence_1;\ldots;\Sequence_n}\DNFEq\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}$
    then
    $n=m$ and
    $\Sequence_i = \SequenceAlt_i$ for all $i\in\RangeIncInc{1}{n}$
  \item
    If $\NormalizeOf{\SequenceOf{\Atom_1;\ldots;\Atom_n}} \DNFEq
    \NormalizeOf{\SequenceOf{\AtomAlt_1;\ldots;\AtomAlt_n}}$ then
    $\NormalizeOf{\Atom_{\sigma_1(i)}} =
    \NormalizeOf{\AtomAlt_{\sigma_2(i)}}$ for all $i$
    where $\sigma_1 =
    \SortingOf{\SequenceLeq}{\ListOf{\NormalizeOf{\Atom_1};\ldots;\NormalizeOf{\Atom_n}}}$
    and $\sigma_2 =
    \SortingOf{\DNFLeq}{\ListOf{\NormalizeOf{\AtomAlt_1};\ldots;\NormalizeOf{\AtomAlt_n}}}$
  \item
    tester
   % If $\NormalizeOf{\StarOf{\DNFRegex}}\DNFEq\NormalizeOf{\StarOf{\DNFRegexAlt}}$
  \end{enumerate}
\end{theorem}

\begin{definition}
  \begin{enumerate}
  \item
    Let
    $\DNFOf{\Sequence_1;\ldots;\Sequence_n}$
    and
    $\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}$
    be two DNF regular expressions, where
    $\NormalizeOf{\DNFOf{\Sequence_1;\ldots;\Sequence_n}} \DNFEq
    \NormalizeOf{\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}}$.
    $\DNFLensSynthOf{\DNFOf{\Sequence_1;\ldots;\Sequence_n}}
    {\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}}=
    (\DNFLensOf{\SequenceLens_1;\ldots;\SequenceLens_n},\sigma)$
    where
    $\SequenceLens_i=\SequenceLensSynthOf{\Sequence_1}{\SequenceAlt_{\sigma(i)}}$
    and
    $\sigma=\SortingOf{\SequenceLeq}{\ListOf{\NormalizeOf{\Sequence_1};\ldots;\NormalizeOf{\Sequence_n}}}
    \Compose
    \SortingOf{\SequenceLeq}{\ListOf{\NormalizeOf{\SequenceAlt_1};\ldots;\NormalizeOf{\SequenceAlt_n}}}$
  \item
    Let
    $\SequenceOf{\String_0;\Atom_1;\ldots;\Atom_n;\String_n}$
    and
    $\SequenceOf{\StringAlt_0;\AtomAlt_1;\ldots;\AtomAlt_n;\StringAlt_n}$
    be two sequences, where
    $\NormalizeOf{\SequenceOf{\String_0;\Atom_1;\ldots;\Atom_n;\String_n}} \SequenceEq
    \NormalizeOf{\SequenceOf{\StringAlt_0;\AtomAlt_1;\ldots;\AtomAlt_n;\StringAlt_n}}$.
    $\SequenceLensSynthOf{\SequenceOf{\String_0;\Atom_1;\ldots;\Atom_n;\String_n}}
    {\SequenceOf{\StringAlt_0;\AtomAlt_1;\ldots;\AtomAlt_n;\StringAlt_n}}=
    (\SequenceLensOf{(\String_0,\StringAlt_0);\AtomLens_1;\ldots;\AtomLens_n;(\String_n,\StringAlt_n)},\sigma)$
    where
    $\AtomLens_i=\AtomLensSynthOf{\Atom_1}{\AtomAlt_{\sigma(i)}}$
    and
    $\sigma=\SortingOf{\AtomLeq}{\ListOf{\NormalizeOf{\Atom_1};\ldots;\NormalizeOf{\Atom_n}}}
    \Compose
    \SortingOf{\SequenceLeq}{\ListOf{\NormalizeOf{\AtomAlt_1};\ldots;\NormalizeOf{\AtomAlt_n}}}$
  \item
    Let
    $\StarOf{\DNFRegex}$
    and
    $\StarOf{\DNFRegexAlt}$
    be two atoms, where
    $\NormalizeOf{\StarOf{\DNFRegex}} \SequenceEq
    \NormalizeOf{\StarOf{\DNFRegexAlt}}$.
    $\AtomLensSynthOf{\StarOf{\DNFRegex}}
    {\StarOf{\DNFRegexAlt}}=
    \IterateLensOf{\DNFLensSynthOf{\DNFRegex}{\DNFRegexAlt}}$
  \end{enumerate}
\end{definition}

\begin{theorem}
  $\NormalizeOf{\DNFRegex}\DNFEq\NormalizeOf{\DNFRegexAlt}$
  if, and only
  if, there exists a $\DNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt$,
  where the typing of $\DNFLens\OfType\DNFRegex\Leftrightarrow\DNFRegexAlt$
  has no applications of a rewrite rule.
\end{theorem}

Unfortunately, while this can find a well typed DNF lens, it doesn't
necessarily find the one that matches the examples.  In the situation where
there are multiple valid sorted orderings, there are multiple different lenses
with potentially different semantics.  Instead of iterating through all the
possible sortings, and finding one that matches the examples, we would like to
be able to immediately find only a sorting which satisfies the examples.

@@

This is done by joining additional parsing information alongside the regular
expression components, and creating an ordering on DNF Regular expressions
with parsing information.\afm{This would be a really in depth thing to describe,
worth going into?}

### 2016-09-16 paper/related-work.tex
added variable information on master file to each file
added rough draft of all sections

\section{Related Work}
\subsection{FlashFill}
\subsection{FlashExtract}
\subsection{SemFill}

>>>>

\section{Related Work}
\subsection{FlashFill}
FlashFill is another string transformation program.  FlashFill takes input and
output examples as its only form of specification.  For FlashFill's primary use
case, providing easy transformations of Excel data, input and output examples as
the only specification makes a lot of sense.  Our tool is intended for use for
programmers to use, who already know regular expressions.  Our tool is oriented
towards a situation where a programmer would check the generated code into a
repository.  From requiring this extra information, our tool is able to handle
synthesize complicated functions that go beyond what FlashFill can synthesize.
For example, given 21 examples of extracting the first author, with the names
reversed, FlashFill still gave incorrect outputs for a number of inputs, even
when those inputs are a similar form to the provided examples.

Furthermore, even if FlashFill were to synthesize the correct program, it would
be foolhardy to check the generated code into a repository.  FlashFill does not
fail on inputs it doesn't know how to handle, rather it tries to provide an
output.  This is different from the generated lenses, which will quickly
fail on inputs that do not match it's intended specification.  Furthermore, the
specification of what inputs should be handled are provided by the user, not
inferred from the examples, giving it more reliability. \afm{talk about how it
  did on bibtex example, more specific on what it did wrong, maybe?}

\subsection{FlashExtract}
FlashExtract is a program that takes an input of a string, and a partial
labeling of substrings within that string, to try to label the rest of the
substrings.  FlashExtract was able to label some amounts of the strings, but was
unable to handle reorderings of the data well.  While it would do well in
extracting data when the data came in the same form, for example, when given a
Bibtex file where the fields come in a specific order, the data would do well,
however, FlashExtract was unable to handle when fields were given in different
orders, even if previous examples in that order were given.

%%% Local Variables:
%%% TeX-master: "main"
%%% End:
